---
title: "tidy-model"
author: "Brandon Emory"
date: "2025-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse') ; library('modeldata')
``` 
## INTRODUCTION 

This R Markdown document demonstrates my abilities to use models for data analysis using a data set collected on crickets.

<!-- Load Data Chunk -->

```{r load data, include = FALSE}
data(crickets, package = "modeldata")
names(crickets)
```

<!-- What factors affect the chirp rate? -->
## What is that sound?

In this report, we examine what factors predict a cricket's chirp rate.

Number of species - 31, temp max 30.4, min 17.2, mean rate of chirping - 72.89

```{r summary, echo = FALSE}
summary(crickets)
ggplot(crickets, aes(x = rate)) +
  geom_histogram(bins = 8) + 
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab('Chirp rate (per min.)')
```

<!-- Graphing temperature and chirps -->
## Temperature affects chirp rate

```{r temp, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')

temp_lm <- lm(rate ~ temp, crickets)

summary.lm(temp_lm)
```
Based on a scatter plot of temperature and chirping and a correlation test, it seems that as temperature increases one degree, the rate of chirping increases about 4.2 chirps per minute.












