---
title: "tidy-model-diamond"
author: "Brandon Emory"
date: "2025-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```  

## Introduction  
This R Markdown document demonstrates my ability to use models for data analysis using a dataset on diamond prices.

```{r load data, include=FALSE}
data(diamonds, package = "ggplot2")
names(diamonds)
```

## What affects diamond prices?  
In this report, we examine what factors predict the price of a diamond.

```{r summary, echo=FALSE}
summary(diamonds)

ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
  ggtitle("Distribution of Diamond Prices") +
  xlab("Price (USD)")
```

## Carat affects price  

```{r carat_price, echo=FALSE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = 'lm') +
  ggtitle("Plot of Carat and Price") +
  xlab("Carat") +
  ylab("Price (USD)")

carat_lm <- lm(price ~ carat, diamonds)
summary.lm(carat_lm)
```

Based on a scatter plot of carat and price, as carat increases, the price also increases. The model suggests that for every additional carat, the price rises significantly.





